Gastrointestinal II

### 2017-09-26 git2.rmd
mc

>>>>

## Trigger 1

DDx:

-  Motility: hyperthyroidism (hyper) or scleroderma (hypo)
-  Osmotic: improves with fasting
-  Exudative: damage to mucosa, e.g. shigella infn
-  Secretory: cholera

Most common cause in his age: infectious; chronic so thinking of some things like parasites, or immunocompromised

Fluids: 8-9L per day, only 1L is from intake whereas rest is from salivary glands,... 
SI = nutrients, LI = water

Hx:

-  HPC: onset, time course (how often, when, has it changed), consistency, how much he produces, how many wet nappies does he create, associated (systemicall unwell, any pain), was there a precipitating illness => antibiotics, who he comes into contact with (indigenous communities are a lot more communal)
-  PMHx: vaccinations

## Trigger 2

Poor diet: can by itself cause diarrhoea, e.g. full fat diet

## Trigger 3

Pale conjunctiva and nail beds: anaemia either from nutrition (Fe or Vit B12) or haemorrhage

Foul smelling, pale soft fluffy stool: malabsorptive. Smelly due to fermentation of the fats in it

Ix:

-  Bloods: FBC, ELFT, cultures (if he was febrile)
-  Stool sample

## Trigger 4

Cysts: not 100% specific or sensitive

Giardia:

-  Flagellated protozoa
-  Fecal-oral transmission
-  Proximal small bowel
-  Incredibly widespread: 70-80% of kids pass cysts but are asymptomatic

Anaemic: iron deficiency is microcytic, B12/folate are macrocytic |

-  Microcytic = {T = thalassaemia, I = iron deficiency, C = chronic disease, S = sideroblastic anaemia}
-  Macrocytic = {A = alcohol (cirrhosis), B = B12 deficiency, C = compensatory reticulocytosis (blood loss and hemolysis), D = drug/dysplasia, E = endocrine (hypothyroidism), F = folate deficiency/foetus (pregnancy)}

## Trigger 5

Metronidazole: inhibits nucleic acid synthesis so it affects both bacteria and protozoa

Long term consequences of parasitic infn:

-  Damage to gut mucosa
-  Can migrate to subarachnoid space => meningitis/encephalitis
-  Malnutrition
-  Failure to thrive
